Space Matters: extending sensitivity analysis to initial spatial conditions in geosimulation models

### 2018-09-03 Response_to_Reviewers.tex

	% 9. The authors should illustrate the contribution of each parameter of the generator to the generated initial pattern using a figure. Details of the generators should be included in an appendix
	\item The behavior of the spatial generator was described with more details in Appendix.

	% 10. Explain how the density grids are used in the Schelling model, do you allow more than one agent to be in the same cell? 
	\item \comment[JR]{il a mal lu ou on est pas assez explicites, a detailler}

	% 11. Line 297: I am not sure why empirical density destitutions of cities are important in the context of this study. 
	\item

	% 12. Line 641: NetLogo supports grids, continuous space, and GIS elements. It does not restrict the user in any way.
    % \comment[JR]{plutot d'accord avec lui mais parce que proficient en NetLogo, peut etre que des debutants sont plus contraints ? quelquun aurait une ref là dessus ?}
	\item We agree that NetLogo does not restrict the use on what implementation to use. However, some features are much more easier as raster or the contrary, what could influence the choice of beginners. We clarified that statement in the paper.

@@

	% You write “To our knowledge there exist no single well established method to compare phase diagrams in the agent-based modelling and geosimulation literature.” As this is new you may want to discuss how say its done in other fields?  
	\item \comment[JR]{cf remarques other reviewers sur ce point}

	% In the Schelling and sugarscape models how are the initial population of agents distributed? Is it random? I ask as this is unclear in the current version of the paper but it is vital if your purpose is to explore how the initial spatial configuration impacts the results of the model. Should the agents not be all located at the same location for each run to rule out the impact of initial placement of agents on the resulting patterns? This is something that is not addressed in the paper. I know for example, the Schelling model in NetLogo normally places the agents at random locations but in your reimplementation how is this done? The same goes for the sugarscape model in NetLogo. I believe a clearer description of the model initialization would be good and how its not the initial random placement of agents but the actual environment that impacts the outcomes. Maybe this was explained in the paper but I missed it.  
	\item \comment[JR]{on regarde des indicateurs aggreges qui ne dependent pas de la trajectoire de chaque agent, si le placement initial est random il na aucune influence sur ceux-ci (cf les distribs) -> ajouter des graphes de comportement stat des modeles en Appendix ?}

	% In Section 1.2 talk about “(monocentric vs. polycentric for example)” but in the results you don’t discuss the monocentric versions explicitly using these terms.
	\item

	% Also I appreciate you are trying to show rather complex results but reading the results several times I found the presentation of them (both written and visual) really hard to follow and understand. I was wondering if maybe you could find a better way to show / discuss them. Maybe a  worked example of how each set of parameters leads to a point in say Figure 4. Also it is confusing why the results from both models are not presented using the same styles of figures. For example, for the sugar scape model you different styles of figures than for the segregation model.   
	\item Additional figures for the Schelling model presenting the results in a similar way were provided. Figures showing grid estimates of phase diagrams could not be realized for the Schelling model because of the different sampling of the parameter space used.
\end{itemize}

### 2018-09-03 paper.tex
Update on Overleaf.

Considerations of such issues in the geographical modelling literature are rather scarce.Despite the opportunity of simulation in the representation of space and its potential observation biases, there have been some noticeable attempts to evaluate the effects of the same set of mechanisms run from different initial spatial conditions, which have focused on the impact of:
\begin{itemize}
\item The accuracy of geo-localised input data;
\item The shape, precision and boundaries of the modelled spatial system;
\item The degree of spatial heterogeneity modelled.
\end{itemize}

@@

An aspect that we have not touched upon in the article with respect to the sensitivity to initial spatial conditions is the importance of the modelling platform as a constraint in the formalisation of space. For example, the use of Netlogo fosters the use of raster/grid spatial inputs rather than vector elements. Its toroidal default setting might also have influenced the work of many modellers who did not question explicitly the representation of space. This issue is part of the docking challenge \citep{Axtelletal1996} (i.e. checking if two models can produce the same results), but more generally, it involves a description of the model and its spatial requirements more detailed than what is currently the rule.

@@

%%%%%%%%%%%%%%
\begin{figure}
	\includegraphics[width=0.49\linewidth]{figures/alpha_beta_examples.png}
    \includegraphics[width=0.49\linewidth]{figures/alpha_beta_schemas.png}
    \caption{\added{\textbf{Stylized behavior of the density grid generator.}}\label{fig:stylized-generator}}
\end{figure}
%%%%%%%%%%%%%%

>>>>

Considerations of such issues in the geographical modelling literature are rather scarce. Despite the opportunity of simulation in the representation of space and its potential observation biases, there have been some noticeable attempts to evaluate the effects of the same set of mechanisms run from different initial spatial conditions, which have focused on the impact of:
\begin{itemize}
\item The accuracy of geo-localised input data;
\item The shape, precision and boundaries of the modelled spatial system;
\item The degree of spatial heterogeneity modelled.
\end{itemize}

@@

An aspect that we have not touched upon in the article with respect to the sensitivity to initial spatial conditions is the importance of the modelling platform as a constraint in the formalisation of space. For example, \replaced{spatial structure may be easier to implement as a raster rather a vector in NetLogo models, what could influence the implementation choices of some non-experienced modelers}{the use of Netlogo fosters the use of raster/grid spatial inputs rather than vector elements}. Its toroidal default setting might also have influenced the work of many modellers who did not question explicitly the representation of space. This issue is part of the docking challenge \citep{Axtelletal1996} (i.e. checking if two models can produce the same results), but more generally, it involves a description of the model and its spatial requirements more detailed than what is currently the rule.

@@

%%%%%%%%%%%%%%
\begin{figure*}
	\includegraphics[width=0.49\linewidth]{figures/alpha_beta_examples.png}
    \includegraphics[width=0.49\linewidth]{figures/alpha_beta_schemas.png}
    \caption{\added{\textbf{Stylized behavior of the density grid generator.}}\label{fig:stylized-generator}}
\end{figure*}
%%%%%%%%%%%%%%
